---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---


# idea
Get statistics for enhancer coverage for different publications

```{r}
library(rtracklayer)
library(ggplot2)
# library(reg2gene)
library(stringr)
library(parallel)
library(GenomicRanges)
library(GenomicInteractions)
library(InteractionSet)
source("/data/akalin/Projects/AAkalin_reg2gene/reg2gene/R/BenchmarkGInteractions.R")
source("/data/akalin/Projects/AAkalin_reg2gene/reg2gene/R/projectHelpFunctions.R")
source("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Scripts/Downstream/plots/generalPlot_functions.R")


# TP from Gaasperini dataset
inhouseM <- readRDS("/data/local/ipatarc/AAkalin_reg2gene/Results/inhouse_models.rds")

enhancer_EGA_paths <- c( "/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/JEME/downloaded/fANTOM_encoderoadmap_lasso_EN_pooled_processed_GR_forceBYname_19_06_19.rds",
                    "/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/GeneHancer/Fishilevich_2017_GeneHancer_hg19_GInteractions_190606.rds",                                                     
                    "/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/EnhancerAtlas/Processed/GInteractions_EnhancerAtlas_ENSG_addedd_20918_forceByName180930.rds",   
                    "/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/Hait_2017_FOCS/processed/19_06_17_FOCS_pooled_processed_interactions.rds")

names(enhancer_EGA_paths) <- c("JEME","GeneHancer","EnhancerAtlas","FOCS")



```





# Analysis
```{r}
     enh_genome_coverage <- getEnhCoverage(enhancer_EGA_paths,names =names(enhancer_EGA_paths) )
    
    # 1. quantify publication info for each enhancer base in the genome
    enh_perBase_perPubl_coverage <- get_perPubl_BaseCoverage(enh_genome_coverage)


```

```{r}
    
    
#######################################################################  
## get number of enhancers per publication
    # IMPORT EGA interactions
        benchData <- lapply(enhancer_EGA_paths,function(x) {print(x)
          return(unique(readRDS(x)))})
        names(benchData) <- names(enhancer_EGA_paths)
        
        
  # extract enhancers and get reduced dataset 
        Enhancers <- lapply(names(benchData),function(x){
          A=reduce(unique(anchorOne(benchData[[x]])))
          A$paper <- x
          return(A)
          
        })
        
            names(Enhancers) <- names(enhancer_EGA_paths)
          
        N_reducedEnhancers <- sapply(Enhancers,length) # after enhancers are reduced together
            
      
        enh_perBase_perPubl_coverage <- sapply(Enhancers, function(x){sum(width(reduce(x)))})
        
################################
# original enhancers
#enhancers defined in publications
     Enhancers_original <- lapply(names(enhancer_EGA_paths),
                                  function(x){
      
             benchData <- unique(readRDS(enhancer_EGA_paths[x]))
      
                  A=unique(anchorOne(benchData))
                  A$paper <- x
                  return(A)
  
                        })
          names(Enhancers_original) <- names(enhancer_EGA_paths)
 
      N_originalEnhancers <- sapply(Enhancers_original,length) # original info about enhancers
    
  
# get info about coverage        
# genome-wide base coverage of enhancers covered by different publications      
  enh <- c(Enhancers$JEME,Enhancers$GeneHancer,
           Enhancers$FOCS,Enhancers$EnhancerAtlas)
  
  Enhancers_coverage <-  sum(width(reduce(enh)))
      
      
# get size of the full genome
    Full.genome <- sum(width(GRanges(coverage(GRangesList(Enhancers)))))    

    
# calculate genome coverage by enhancer definition      
       EnhGenome_coverage <- round(100*enh_perBase_perPubl_coverage/Full.genome,1)
       EnhEnh_coverage <- round(100*enh_perBase_perPubl_coverage/Enhancers_coverage,1)
       
       
       print(paste0("percentage of the genome covered by enhancers: ",EnhGenome_coverage))
       
       
      
enh_stat <- data.frame(cbind(EnhGenome_coverage,
                             EnhEnh_coverage,
                  N_reducedEnhancers[names(EnhGenome_coverage)],
                  N_originalEnhancers[names(EnhGenome_coverage)]))
colnames(enh_stat) <- c("EnhGenome_coverage",
                        "EnhEnh_coverage",
                        "N_reducedEnhancers",
                        "N_originalEnhancers")

###############################
# PLOT THE DATA USING GGPLOT2


         colorScheme <- c("black","darkorange","blue","darkcyan")
            names(colorScheme) <- c("FOCS","GeneHancer","JEME","EnhancerAtlas")
        
        path="/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Results/Downstream/Plots/figure1_publicationE/fig1A_bubbleplot_enh_def.png"
        
        png(path,width = 900, height = 900)
          
          
          ggplot(data=enh_stat, aes(x=N_reducedEnhancers, y=EnhGenome_coverage)) +
            geom_point(aes(size=N_originalEnhancers,
                           colour=rownames(enh_stat))) +
            scale_color_manual(values=colorScheme[rownames(enh_stat)])+
            scale_size_continuous(range=c(10,40)) +
            theme(legend.position = "none",
                  plot.margin = unit(c(1,1,2.5,2.5), "cm"),
                  panel.grid.major = element_blank(), 
                  panel.grid.minor = element_blank(),
                  panel.background = element_blank(), 
                  axis.line = element_line(colour = "black"),
                  axis.title = element_text(size=50,face="bold"),
                  axis.text =element_text(size=40,face="bold"),
                  text = element_text(size=40,face="bold", vjust = 2))+
            geom_text(aes(label=rownames(enh_stat)),
                      hjust=0.5, vjust=-2,size=15, fontface="bold")+
            xlab("\nN of enhancers") +
            ylab("% of genome coverage\n") +
            scale_x_continuous(limits = c(0, 350000),labels = scales::comma)+
            scale_y_continuous(limits = c(0, 40))
          
        
          dev.off()

      
#############################################################
# print table for statistics
          
          
           png("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Results/Downstream/Plots/fig1A_bubbleplot_enh_def_inputdata.png" , 
               height=280, width=600)
           library(gridExtra)
           p<-tableGrob(enh_stat)
           grid.arrange(p)
           dev.off()
            
```

#n of basepairs covered by combiantion of enhancer definitions: 1-4

```{r}
             

# n of basepairs covered by combiantion of enhancer definitions: 1-4
        
coverage <- list() 
   
  for (i in 1:4){
      coverage[[i]] <- sum(width(enh_genome_coverage[enh_genome_coverage$score==i]))
    }

    # pooling data
      enh_covered_by_methods <- data.frame(round(100*unlist(coverage)/Enhancers_coverage,1))

    enh_covered_by_methods <- cbind(enh_covered_by_methods,rownames(enh_covered_by_methods))
     colnames(enh_covered_by_methods) <- c("value","category")
     

path="/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Results/Downstream/Plots/supp_fig1A_enhCoverage.png"
png(path,width = 900, height = 800)



 ggplot(enh_covered_by_methods, aes(y=value, x=category)) + 
    geom_bar(position="dodge", stat="identity",fill="orange")+
  scale_y_continuous(limits = c(0, 100))+
    geom_text(aes(label=value), vjust=0,size=15) +
 ylab("%[enhancer def coverage]\n") +
    xlab("\nEnhancer category") +
    theme(#legend.position = "none",
          plot.margin = unit(c(1,1,4,2.5), "cm"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black"),
          text = element_text(size=40,face="bold"))
 
 
 dev.off()


```



Results: 
925479036 bases in the genome is covered by one definition of enhancer 72.4%
311707159 bases in the genome is covered by two definition of enhancer 24.4%
37735540 bases in the genome is covered by three definition of enhancer 2.9%
3760226 bases in the genome is covered by four definition of enhancer 0.3%

