---
title: "Creating set of negatives for other publications and obtaining statistics"
output: html_document
---


#IDEA: 

Check how enhancer regions are analyzed by ChromHMM. If they correspond to enhancers,promoters, etc.


#HOW TO

1) individually for all FOCS/JEME/GeneHancer/EnhancerAtlas/inhouse enhancers screen trough ChromHMM predictions and identify if some regions are repressed, heterochromatin, etc across all cell types (get mnemomics)*

2) get statistics for enhancer regions that are actually repressed

3) overlap repressed enhancers and identify in how many EP links they apprear


__*Create a set of negatives that will be used to test our models__
Negatives will be selected from the set of enhancers that we are testing in by individual published method, by screening these regions throuhg ChromHMM models across all Roadmap cell types, 
searching for reg2gene enhancer regions that were predicted by ChromHMM as __NOT__ being "12_EnhBiv","11_BivFlnk","10_TssBiv","8_ZNF/Rpts","7_Enh","6_EnhG","5_TxWk","4_Tx","2_TssAFlnk","3_TxFlnk","1_TssA" in __ALL__ cell types. In other words, enhancer regions that are repressed (quies, polycomb repressed, etc.) or similar across all Roadmap cell types 



step1 compare individual publications enhancer region definition (confirming we do good job with enhancers, MARCH 2019)
step 2 identify EGA that are based on enhancer regions which should be repressed


__help__
trello board with described mnemomics:
for reg2gene
https://trello.com/c/O02fe5Ur/303-performance-assessment-creating-negatives
for other methods
https://trello.com/c/lnTXqz1z/320-proof-of-concept-enhancer-regions


```{r setup, include=FALSE}

# "/home/ipatarc/R/x86_64-redhat-linux-gnu-library/3.5"


```

## STEP 1: Get mnemonics
 IDEA: detecting if in FOCS/JEME/GeneHancer/EnhancerAtlas set of enhancers I can identify enhancers that are repressed in the full dataset
 
```{r , eval=FALSE}

.libPaths("/home/ipatarc/R/x86_64-redhat-linux-gnu-library/3.5")
library(GenomicInteractions)
library(stringr)
library(pheatmap)
source("/data/akalin/Projects/AAkalin_reg2gene/reg2gene/R/BenchmarkGInteractions.R")
source("/data/akalin/Projects/AAkalin_reg2gene/reg2gene/R/projectHelpFunctions.R")

image.path <- "/data/akalin/Projects/AAkalin_Catalog_RI/AAkalin_CatalogRI/Results/Plots/"


# mnemonics files input
mnemonicsFil <- list.files("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/Roadmap/All_mnemonics_bedFiles", pattern = ".bed$", full.names = T)





names <- c("inhouseModels","flexible","stringent","jeme","GeneHancer","EnahncerAtlas","FOCS")
Publ = c("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Results/Inhouse_Enhancers_and_EGA_Analysis/EGA/step4_filtering_and_voting/voting/inhouse_models.rds", 
           "/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Results/Inhouse_Enhancers_and_EGA_Analysis/EGA/step5_unifyingEnhDef/flexible.rds",
           "/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Results/Inhouse_Enhancers_and_EGA_Analysis/EGA/step5_unifyingEnhDef/stringent.rds",
           "/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/JEME/downloaded/fANTOM_encoderoadmap_lasso_EN_pooled_processed_GR_forceBYname_19_06_19.rds",
           "/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/GeneHancer/Fishilevich_2017_GeneHancer_hg19_GInteractions_190606.rds",
           "/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/EnhancerAtlas/Processed/GInteractions_EnhancerAtlas_ENSG_addedd_20918_forceByName180930.rds",
           "/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/Hait_2017_FOCS/processed/19_06_17_FOCS_pooled_processed_interactions.rds")
           
 names(Publ) <- names     


  path1="/data/local/ipatarc/AAkalin_CatalogRI/Data/mnemonics/"


 lapply(Publ,
              function(x){
                # x=Publ[5]
                
                print(x)
                interactions = readRDS(x)
                
               
                enhancerRegions=unique(anchorOne(interactions))
                      print(paste0("N enhancers",length(enhancerRegions)))
                # Filtering out chrY
                enhancerRegions <- enhancerRegions[!seqnames(enhancerRegions)%in%c("chrY","chrM")]
                      print(paste0("N enhancers without chrY",length(enhancerRegions)))
      
                      # enhancerRegions=enhancerRegions[1:1000]
                # splitting data into chunks
                 
      Split.factor <- split(1:length(enhancerRegions),
                      sort(1:length(enhancerRegions)%%as.integer(100)))

      print(length(Split.factor))
     
       
      Split.Results = lapply(Split.factor,function(split.round){
    
      
            print(unlist(split.round))
       
                Mnem_FOR_Enh <- getMnemonics(mnemonicsFil,
                                             enhancerRegions=enhancerRegions[unlist(split.round)],
                                             mc.cores=30)
              
                saveRDS(Mnem_FOR_Enh,
                    paste0(path1,
                           "chunk",
                           split.round[[1]][1],
                           str_replace(basename(x),".rds",""),"RegionsMnemonics_19_01_23.rds"))
               
              })


      
      # Pooliing data together
      
            mnemLists <- lapply(list.files(path1,
                                           pattern=str_replace(basename(x),".rds",""),full.names = T),
                   readRDS)
      
                       mnemListsPooled <- do.call("c",mnemLists)
                       mnemListsPooled <- lapply(mnemListsPooled,as.character)
                       mnemLists.df <- do.call("rbind.data.frame",mnemListsPooled)
                   
                  rownames(mnemLists.df) <- names(mnemListsPooled)
                  colnames(mnemLists.df) <- rownames(mnemLists[[1]])
             
                                
               saveRDS(mnemLists.df,paste0(path1,"Full",
                                 str_replace(basename(x),".rds",""),"RegionsMnemonics_19_01_23.rds"))
               
            
               
               
                })
 
 
            system("cp $(find . | grep Full) /data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Results/Mnemomics/Enhancers/")              
          
```

===========================================
waiting for results

# How many FOCS interactions should be reported as repressed?
n[EGA]=302
n[e]=
```{r }

library(GenomicInteractions)
library(stringr)
library(pheatmap)
source("/data/akalin/Projects/AAkalin_reg2gene/reg2gene/R/BenchmarkGInteractions.R")
source("/data/akalin/Projects/AAkalin_reg2gene/reg2gene/R/projectHelpFunctions.R")

Mnem_FOR_Enh <- readRDS("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Results/Mnemomics/Enhancers/Full19_06_17_FOCS_pooled_processed_interactionsRegionsMnemonics_19_01_23.rds")


# creating Granges object
      # Mnem_FOR_Enh_2 <- t(Mnem_FOR_Enh)
      
      EnhRegions <- GRanges(str_replace(str_replace(rownames(Mnem_FOR_Enh),"\\.",":"),"\\.","-"))
            mcols(EnhRegions) <- DataFrame(Mnem_FOR_Enh)


# binarize mnemonics marks, count and report in how many cell types this region is repressed 
perEnhancerRepressedBinary <- binarizeFunction(EnhRegions,
                                               categories= c("12_EnhBiv","11_BivFlnk","10_TssBiv","8_ZNF/Rpts","7_Enh","6_EnhG","5_TxWk","4_Tx","2_TssAFlnk","3_TxFlnk","1_TssA" ))



RepressedEnhRegions <- EnhRegions[which(perEnhancerRepressedBinary==127)]



FOCS_summary <- (length(RepressedEnhRegions))
```




# How many JEME interactions should be reported as repressed?
671
```{r }


library(GenomicInteractions)
library(stringr)

Mnem_FOR_Enh <- readRDS("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Results/Mnemomics/Enhancers/FullfANTOM_encoderoadmap_lasso_EN_pooled_processed_GR_forceBYname_19_06_19RegionsMnemonics_19_01_23.rds")


# creating Granges object
      # Mnem_FOR_Enh_2 <- t(Mnem_FOR_Enh)
      
      EnhRegions <- GRanges(str_replace(str_replace(rownames(Mnem_FOR_Enh),"\\.",":"),"\\.","-"))
            mcols(EnhRegions) <- DataFrame(Mnem_FOR_Enh)


# binarize mnemonics marks, count and report in how many cell types this region is repressed 
perEnhancerRepressedBinary <- binarizeFunction(EnhRegions,
                                               categories= c("12_EnhBiv","11_BivFlnk","10_TssBiv",
                                                             "8_ZNF/Rpts",
                                                             "7_Enh","6_EnhG",
                                                             "5_TxWk","4_Tx","2_TssAFlnk","3_TxFlnk","1_TssA" ))

RepressedEnhRegions <- EnhRegions[which(perEnhancerRepressedBinary==127)]
JEME_summary <- (length(RepressedEnhRegions))

```






# How many GeneHancer interactions should be reported as repressed?

```{r }


library(GenomicInteractions)
library(stringr)

Mnem_FOR_Enh <- readRDS("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Results/Mnemomics/Enhancers/FullFishilevich_2017_GeneHancer_hg19_GInteractions_190606RegionsMnemonics_19_01_23.rds")


# creating Granges object
      # Mnem_FOR_Enh_2 <- t(Mnem_FOR_Enh)
      
      EnhRegions <- GRanges(str_replace(str_replace(rownames(Mnem_FOR_Enh),"\\.",":"),"\\.","-"))
            mcols(EnhRegions) <- DataFrame(Mnem_FOR_Enh)


# binarize mnemonics marks, count and report in how many cell types this region is repressed 
perEnhancerRepressedBinary <- binarizeFunction(EnhRegions,
                                               categories= c("12_EnhBiv","11_BivFlnk","10_TssBiv",
                                                             "8_ZNF/Rpts",
                                                             "7_Enh","6_EnhG",
                                                             "5_TxWk","4_Tx","2_TssAFlnk","3_TxFlnk","1_TssA" ))

RepressedEnhRegions <- EnhRegions[which(perEnhancerRepressedBinary==127)]
GH_summary <- (length(RepressedEnhRegions))

```



# How many EnhancerAtlas interactions should be reported as repressed?

```{r }


library(GenomicInteractions)
library(stringr)

Mnem_FOR_Enh <- readRDS("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Results/Mnemomics/Enhancers/FullGInteractions_EnhancerAtlas_ENSG_addedd_20918_forceByName180930RegionsMnemonics_19_01_23.rds")


# creating Granges object
      # Mnem_FOR_Enh_2 <- t(Mnem_FOR_Enh)
      
      EnhRegions <- GRanges(str_replace(str_replace(rownames(Mnem_FOR_Enh),"\\.",":"),"\\.","-"))
            mcols(EnhRegions) <- DataFrame(Mnem_FOR_Enh)


# binarize mnemonics marks, count and report in how many cell types this region is repressed 
perEnhancerRepressedBinary <- binarizeFunction(EnhRegions,
                                               categories= c("12_EnhBiv","11_BivFlnk","10_TssBiv",
                                                             "8_ZNF/Rpts",
                                                             "7_Enh","6_EnhG",
                                                             "5_TxWk","4_Tx","2_TssAFlnk","3_TxFlnk","1_TssA" ))


RepressedEnhRegions <- EnhRegions[which(perEnhancerRepressedBinary==127)]
EA_summary <- length(RepressedEnhRegions)
```










```{r, results="asis" }


full_summary <-as.data.frame( rbind(EA_summary,FOCS_summary,GH_summary,JEME_summary))
full_summary$Publication <- as.factor(c("EnhancerAtlas","FOCS","GeneHancer","JEME"))


#############################
# version 2 plot


 colorScheme <- c("black","darkorange","blue","darkcyan" )
    names(colorScheme) <- c("FOCS","GeneHancer","JEME","EnhancerAtlas")
colors <- colorScheme[levels(full_summary$Publication)]


png("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Results/Downstream/Plots/figure1_publicationE/fig1H_repressedEnhancers.png",
    width = 1200,height = 700)

ggplot(full_summary,aes(x=Publication,y=V1,fill=Publication)) +  
  stat_summary(fun.y=mean,position="dodge",geom="bar",width = 0.5) +  
  scale_fill_manual(values =  alpha(colors, .7))+
  
  ylab("N repressed enhancers \n in Roadmap dataset\n ") +
  xlab("\nPublication") + 
  
  theme_minimal()+
  theme(text = element_text(size=40,face="bold"),
        axis.title=element_text(size=50,face="bold"),
        axis.text =element_text(size=40,face="bold"),
        #legend.text = element_text(size=30,face="bold"),
        legend.position="none")+
  geom_text(aes(label = V1), size = 16, hjust = 0.50, vjust = -0.65, position = "stack")+
  ylim(0,15000)


dev.off()

```