---
title: "R Notebook"
output:
html_document:
df_print: paged
---

# IDEA: Data download and processing log for GeneHancer

# STEPS
Download: Y
Pooling: N
Data Organization: Y
Mergining downloaded info: N
Liftovering: Y
Filtering: Y
Forcing coordinates:Y


##Download: Y


FOLDER: /data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/GeneHancer/



# paper summary:
From https://genecards.weizmann.ac.il/geneloc/index.shtml, we downloaded single  .csv document (GeneHancer enhancers are not cell-type specific) with 243,281 GeneHancer enhancers on the 28.11.17. GeneHancer version was 4.6 and it contained enhancer locations and corresponding metadata; including names of associated genes. We wrote an inhouse R scripts to: 1) extract locations of enhancers, 2) obtain one-to-one mapping between genes and enhancers (necessary since more than one gene was reported to be associated with single enhancer region saved as one row of the table), 3) liftover hg38 to hg19 enhancer coordinates (liftOver and rtracklayer R packages), 4) obtain gene locations either by using GeneHancer-reported HCGN gene names or ENSEMBL IDs (biomaRt package, for the genome version GRCh37); and 5) set gene coordinates to be equal to the extended TSS coordinates (+/-1000 bp) of the GENCODE v24 genes.  Post processing, we defined 506,471 unique GeneHancer unique enhancer-promoter association



# Pooling:

__1. input and functions__

```{r, eval=FALSE}

GeneHancer <- read.csv("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/GeneHancer/Fishilevich_2017_GeneHancer_Database/genehancer.csv")



####################
# FUNCTIONS

# extracting genes names

library(stringr)
library(parallel)

    extract_genes <- function(x){
        genes <- str_extract(x,"connected_gene=.*")
        genes <- str_replace_all(genes,"score=[0-9]*.[0-9]*|connected_gene=","")
        genes <- unlist(str_split(genes,";"))
        genes <- genes[nchar(genes)>0]
    }
    
# extracting enhancers names
      extract_scores <- function(x){
        
        genes <- unlist(str_extract_all(x,"score=[0-9]*.[0-9]"))
        genes <- str_replace_all(genes,"score=","")
       
      }

# create 1-to-1 datatable (one enh-1 gene)
      oneToOne <- function(x,
                           GeneHancer,
                           GenesPerEnh,
                           ScoreGenesPerEnh){
        
        df <- GeneHancer[x,]
        df <- df[rep(seq_len(nrow(df)),
                     length( GenesPerEnh[[x]])),]
        df$genes <- GenesPerEnh[[x]]
        df$geneScore <- ScoreGenesPerEnh[[x]]
        return(df)
}



 
 
```


 STEP 0:          
## get one-to-one mappig between genes and enhancers
```{r, eval=FALSE}

#####################
# ANALYSIS
  
      
# STEP 0:          
# get promoters and their scores      
GenesPerEnh <- mclapply(GeneHancer$attributes,extract_genes,mc.cores = 15)
ScoreGenesPerEnh <- mclapply(GeneHancer$attributes,
                             extract_scores,
                             mc.cores = 30)

# create 1-to-1 datatable (one enh-1 gene)
        CreateGeneHancer <- mclapply(1:nrow(GeneHancer),
                                   #1:10,
                                   oneToOne,
                                   GenesPerEnh=GenesPerEnh,
                                   ScoreGenesPerEnh=ScoreGenesPerEnh,
                                   GeneHancer=GeneHancer,
                                   mc.cores = 30)
      
          CreateGeneHancer <- list()
        
       for (i in 1:nrow(GeneHancer)){
                                     #1:10,
                                    print(i)
         CreateGeneHancer[[i]] <-     (oneToOne(x=i, 
                                          GeneHancer=GeneHancer,
                                          ScoreGenesPerEnh=ScoreGenesPerEnh,
                                          GenesPerEnh=GenesPerEnh))
       }
  CreateGeneHancer <- do.call("rbind.data.frame",CreateGeneHancer)

  
  saveRDS(CreateGeneHancer,"/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/GeneHancer/Fishilevich_2017_GeneHancer_Database/18_06_05_GenEnhancerhg38_oneE_oneGenefor.rds")      
  


```



## get GRCH37 coordinates for genes
  
```{r, eval=FALSE}

# STEP 2
# liftover Enhancer coordinates
      GeneHancerFull <- readRDS("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/GeneHancer/18_06_05_GenEnhancerhg38_oneE_oneGenefor.rds")
library(rtracklayer)

library(InteractionSet)
ch = import.chain("/data/akalin/Base/Annotation/hg19/Liftover/hg38ToHg19.over.chain")

# create GRamges object
  GeneHancer.gr <- GRanges(as.character(GeneHancerFull[,1]),IRanges(as.integer(as.character(GeneHancerFull[,4])),as.integer(as.character(GeneHancerFull[,5]))))
  colnames(GeneHancerFull) <- paste("C",colnames(GeneHancerFull),sep = "_")
  mcols(GeneHancer.gr) <- GeneHancerFull
  
  seqlevelsStyle(GeneHancer.gr) = "UCSC"  # necessary

  # liftover  
  cur19.lo <- liftOver(GeneHancer.gr, ch)
 
  # perform merging in 500 chunks cause otherwise to slow

  
  chunks=1000
  Split.factor <- split(1:length(cur19.lo),
                        sort(1:length(cur19.lo)%%chunks))
  
  
  Chunks_merge <- function(x,cur19.lo){
    
    cur19 = do.call("c",cur19.lo[x])
    saveRDS(cur19,
            paste0("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/GeneHancer/EP/",x[1],"chunk.rds"))
             
             
  }
  
 
  parallel::mclapply(Split.factor,function(x){
  
    Chunks_merge(x,cur19.lo=cur19.lo)
    
  },
                       mc.cores = 25)
   
    
  
  # Pooling togethre
        
        allGH <- list.files("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/GeneHancer/EP/", pattern="chunk.rds",full.names = T)
        
        
        ALLdATA <- lapply(allGH,readRDS)
        cur19 = do.call("c",ALLdATA)
        
       saveRDS(cur19,
               "/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/GeneHancer/GeneHancerhg19_EPlinksGRanges.rds") 



```

# getting gene coordinates
+ filtering chrY
```{r, eval=FALSE}

# creating GInteractions obj
      library(InteractionSet)

    cur19 <- readRDS("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/GeneHancer/GeneHancerhg19_EPlinksGRanges.rds") 



# getting gene coordinates by matching names with GENCODE
      tss <- readRDS("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/GENCODE/GENECODE_TSSes.rds")
      
      
      # matching by ensembl id & by common name
          gene.name.matching <- cbind(match(cur19$C_genes,tss$gene.id),
                                      1:length(cur19))
          gene.name.matching <- gene.name.matching[complete.cases(gene.name.matching),]
      
      # matching by ensembl id & by common name
          gene.name.matching2 <- cbind(match(cur19$C_genes,tss$gene.name),
                                       1:length(cur19))
          gene.name.matching2 <- gene.name.matching2[complete.cases(gene.name.matching2),]
          
        gene.name.matching <- rbind(gene.name.matching,gene.name.matching2)
          
      cur19.SS <- GInteractions(cur19[gene.name.matching[,2],],
                               (tss[gene.name.matching[,1],]))
      
        #add metadata 
          mcols(cur19.SS) <- mcols(cur19[gene.name.matching[,2],])
      
          cur19.SS$name <- tss$gene.id[gene.name.matching[,1]]
          cur19.SS$name2 <- tss$gene.name[gene.name.matching[,1]]

     # filtering for chrY     
          cur19.SS <- cur19.SS[seqnames(anchorOne(cur19.SS))!="chrY"]

saveRDS(unique(cur19.SS),
        "/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/GeneHancer/Fishilevich_2017_GeneHancer_hg19_GInteractions_190606.rds")
    
```


# getting enhancers
```{r, eval=FALSE}


# STEP 2
# liftover Enhancer coordinates
      GeneHancer.gr <- readRDS("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/GeneHancer/GeneHancerDatahg38ORGINAL.rds")
library(rtracklayer)
#library(liftOver)
library(InteractionSet)
ch = import.chain("/data/akalin/Base/Annotation/hg19/Liftover/hg38ToHg19.over.chain")

# create GRamges object
  
  seqlevelsStyle(GeneHancer.gr) = "UCSC"  # necessary

  # liftover  
  cur19.lo <- liftOver(GeneHancer.gr, ch)
 
  # perform merging in 500 chunks cause otherwise to slow

  
  chunks=500
  Split.factor <- split(1:length(cur19.lo),
                        sort(1:length(cur19.lo)%%chunks))
  
  
  Chunks_merge <- function(x,cur19.lo){
    
    cur19 = do.call("c",cur19.lo[x])
    saveRDS(cur19,
            paste0("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/GeneHancer/TMP/",x[1],"chunk.rds"))
             
             
  }
  
 
  parallel::mclapply(Split.factor,function(x){
  
    Chunks_merge(x,cur19.lo=cur19.lo)
    
  }, mc.cores = 25)
   
    
  
  # Pooling togethre
        
          allGH <- list.files("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/GeneHancer/TMP/", pattern="chunk.rds",full.names = T)
        
        
        ALLdATA <- lapply(allGH,readRDS)
        cur19 = do.call("c",ALLdATA)
        
       saveRDS(cur19,
               "/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/GeneHancer/GeneHancerhg19_Enhancers.rds") 
 
```



# separate processing of Table4
```{r, eval=FALSE}

# script to process Fishillevic data set:
    # 1) creating GInteractions dataset
    # 2) adding ensembl names  
    # 3) filtering genes that have identified names
    # 4) perform forcing by name

# extracting enhancers from Fish

Fishilevich <- readRDS("/data/akalin/Base/AccessoryData/Fishilevich_2017_GeneHancer_Database/Fishilevich_2017_GeneHancer_Database_GR.rds")
FishE <- Fishilevich[Fishilevich$C_type=="enhancer"]


# creating GInteractions dataset
library(GenomicInteractions)

FishillevicEnhGIx <- GInteractions(FishE,FishE$reg)
    
    mcols(FishillevicEnhGIx) <- mcols(FishE)



source("/data/akalin/Projects/AAkalin_reg2gene/reg2gene/R/BenchmarkGInteractions.R")

    
  # adding ensembl names  
    
    library('biomaRt')
    library(dplyr)
    
    mart <- useDataset("hsapiens_gene_ensembl", useMart("ensembl"))
    genes <- FishillevicEnhGIx$C_targets
        G_list <- getBM(filters= "hgnc_symbol", attributes= c("ensembl_gene_id","hgnc_symbol"),values=genes,mart= mart)
        G_list <- G_list[!duplicated(G_list$hgnc_symbol),] #remove duplicated genes
    
    # filtering genes that have identified names
    
        matchedRegions <- (match(G_list$hgnc_symbol,FishillevicEnhGIx$C_targets))
    
        FishillevicEnhGIx <- FishillevicEnhGIx[matchedRegions]
              FishillevicEnhGIx$name <- G_list$ensembl_gene_id
              FishillevicEnhGIx$name2 <- G_list$hgnc_symbol
          
              
  # perform forcing by name
              
  interactions <- readRDS("/data/local/ipatarc/AAkalin_reg2gene/Results/reg2geneLinks_ModelRes_18_10_01.rds")
    
    FishForced <- forceByname(benchInteractions = FishillevicEnhGIx,interactions = interactions)    

    
   
    saveRDS(FishForced,"/data/akalin/Base/AccessoryData/Fishilevich_2017_GeneHancer_Database/Fishilevich_2017_Table4Enhancers_GeneHancer_forceByName.rds")
    
 
```


# STATISTICS
1. Number of interactions
2. Number of enhancers
```{r, echo=F, message=F }
 length(readRDS("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/GeneHancer/Fishilevich_2017_GeneHancer_hg19_GInteractions_190606.rds"))
length(readRDS("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/GeneHancer/GeneHancerhg19_Enhancers.rds"))
```



```{r}

sessionInfo()
```
