---
title: "R Notebook"
output:
html_document:
df_print: paged
---

# IDEA: Data download and processing log for EnhancerAtlas

# STEPS
Download: Y
Pooling: Y
Data Organization: Y
Mergining downloaded info: Y
Liftovering: N
Filtering: N
Forcing: Y


##Download: Y

# paper summary:

Downloaded on 18.11.2020. from ftp://ftp.ebi.ac.uk/pub/databases/ensembl/encode/integration_data_jan2011/byDataType/openchrom/jan2 011/dhs_gene_connectivity/ genomewideCorrs_above0.7_promoterPlusMinus500kb_withGeneNames_32celltypeCategories.bed8.gz



__1. Creating EG pairs from download __

```{r, eval=FALSE}


# Short processing of EnhancerAtlas


# DATA INPUT
.libPaths("~/R/x86_64-redhat-linux-gnu-library/3.5/")
library(stringr)
library(GenomicRanges)
library(InteractionSet)

# function that uses coordinates from GENCODE and forces EGA object to have equal coordinates
forceByname <- function(benchInteractions,interactions){
  
  # identifying gene names present in benchmark object and object used to 
  # benchmark
  
  require(dplyr)
  
  # test if gene info present in data, if not forceByName should not be used
  if (!any(str_detect(colnames(mcols(benchInteractions)),"name"))) {
    
    return(benchInteractions)
    print("Non gene meta-data present in benchmark dataset")
    
  }
  
  if (!any(str_detect(colnames(mcols(interactions)),"gene.name"))) {
    
    return(benchInteractions)
    print("Non gene meta-data present in interactions dataset")
    
  }
  
  
  # name needs to be found in both datasets to run forceByName argument
  
  if (all(any(str_detect(colnames(mcols(benchInteractions)),"name"))&
          any(str_detect(colnames(mcols(interactions)),"gene.name")))) {
    
    # merging in the manner of findOverlaps
    AllBenchmarkEntries <- 1:length(benchInteractions)
    
    df1 <- data.frame(cbind(AllBenchmarkEntries,
                            as.character(benchInteractions$name)))
    df2 <- data.frame(cbind(1:length(interactions),
                            interactions$gene.name))
    colnames(df1) <- colnames(df2) <- c("integer","gene")
    
    # overlapping left_join  
    matchedPairs <- left_join(df1,df2,by="gene")
    
    # removing duplicated entries
    matchedPairs <- matchedPairs[!duplicated(matchedPairs[,1]),]
    newLocations <- which(!is.na(matchedPairs[,3])) # pairs which name can be forced
    oldLocations <- which(is.na(matchedPairs[,3])) # pairs which name cannot be forced
    
    # matched pairs - pooling together
    
    queryHits <- as.integer(as.character(matchedPairs[newLocations,3]))
    subjetHits <- as.integer(as.character(matchedPairs[newLocations,1]))
    
    # creating GInteractions object    
    benchInteractionsGenes <- interactions[queryHits]
    benchInteractionsMatch <- benchInteractions[subjetHits]
    benchInteractionsNOMatch <- benchInteractions[as.integer(as.character(matchedPairs[oldLocations,1]))]
    
    # pooling info together of those that can be forced by name and those that cannot  
    BenchmarkDSForcedName=GInteractions(anchor1 = anchorOne(benchInteractionsMatch),
                                        anchor2 = benchInteractionsGenes)
    mcols(BenchmarkDSForcedName) <- mcols(benchInteractionsMatch)
    
    BenchmarkDSForcedName <- c(benchInteractionsNOMatch,
                               BenchmarkDSForcedName)
    
    
    
    return(BenchmarkDSForcedName)
  }
}

Thurman <- read.table("/data/akalin/Projects/AAkalin_Catalog_RI/post_PhD_analyses/Data/Thurman2012/genomewideCorrs_above0.7_promoterPlusMinus500kb_withGeneNames_32celltypeCategories.bed8", sep="\t",fill = T,he=F)


```

# data organzation EP links:
1. getting GInteractions object

```{r, eval=FALSE}

library(InteractionSet)


      
    Thurman.GI <- GInteractions(GRanges(Thurman[,1],IRanges(Thurman[,2],Thurman[,3])),
                  (GRanges(Thurman[,5],IRanges(Thurman[,6],Thurman[,7]))))

        Thurman.GI$name=as.character(Thurman[,4])

```



# force to be gencode

```{r, eval=FALSE}
# import source code for forcebyname()


interactions <- readRDS("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/GENCODE/GENECODE_TSSes.rds")


# force TSS coorinates
BTH.GI.f <- forceByname(benchInteractions = Thurman.GI,
                        interactions = interactions)

saveRDS(unique(BTH.GI.f),"/data/akalin/Projects/AAkalin_Catalog_RI/post_PhD_analyses/Data/Thurman2012/Thurman.GI.forcedByName.rds")

```

```{r}

sessionInfo()
```
