---
title: "R Notebook"
output:
html_document:
df_print: paged
---

# IDEA: Data download and processing log for CT FOCS

# STEPS
Download: Y
Pooling: Y
Data Organization: Y
Mergining downloaded info: N
Liftovering: N
Filtering: N


##Download: Y
SOURCE: http://acgt.cs.tau.ac.il/ct-focs/download.html

# paper summary:
We accessed CTFOCS enhancers (on the 21.11.2020.) 

# Data Organization:

__1. Creating EG pairs from download 2__

```{r, eval=FALSE}


library(GenomicInteractions)
library(stringr)
library(InteractionSet)
library(parallel)


setwd("/data/akalin/Projects/AAkalin_Catalog_RI/post_PhD_analyses/Data/CT-FOCS/")
  names <- c("encode","roadmap","fantom")

  FOCS_encode <- read.table("encode.cell_type_specific_ep_links.txt",he=T, fill=T,stringsAsFactors = F)

  FOCS_roadmap <- read.table("roadmap.cell_type_specific_ep_links.txt",he=T, fill=T,stringsAsFactors = F,sep="\t")
  
  
  
  
 Input_Interactions <- c(nrow(FOCS_encode),
                           nrow(FOCS_roadmap))
  
 names(Input_Interactions) <- c("ENCODE",
                                "Roadmap") 
  
 print("Input N of interactions")
 print(Input_Interactions)
 
 
 FOCS <- rbind(FOCS_encode,
        FOCS_roadmap)
 
 
```
# Separate processing for FANTOM5 data

```{r, eval=FALSE}
TMP <- readRDS("/data/akalin/Projects/AAkalin_Catalog_RI/post_PhD_analyses/Data/CT-FOCS/fantom5.EP_links.cell.rds")
tmp <- unlist(TMP)

FOCS_data <- cbind(as.character(names(tmp)),tmp)

info <- str_split(FOCS_data[,1],"\\$")


chr <- sapply(info,function(x){ x <- unlist(x)[3]})
loc <- sapply(info,function(x){ x <- unlist(x)[4]})
name <- sapply(info,function(x){ x <- unlist(x)[2]})
 

gene.loc <- GRanges(chr,IRanges(as.integer(loc),
                                as.integer(loc)))

enh.object <- GRanges(FOCS_data[,2])

CT_FOCS_fantom <- GInteractions(enh.object,gene.loc)
CT_FOCS_fantom$gene.name <- name

```


data organzation EP links:
1. getting GInteractions object
2. remove chrY
3. getting gene coordinates (since only promoter coordinates reported)
```{r, eval=FALSE}

prom <- GRanges(FOCS[,1],IRanges(FOCS[,2],FOCS[,3]),FOCS[,4])
enh <- GRanges(FOCS[,5],IRanges(FOCS[,6],FOCS[,7]))

FOCS_GI <- GInteractions(enh,prom)
mcols(FOCS_GI) <- FOCS

tmp <- FOCS_GI$sym_id
mcols(FOCS_GI) <- NULL
FOCS_GI$gene.name <- tmp

CT_FOCS <- c(CT_FOCS_fantom,FOCS_GI)


# remove chrY
CT_FOCS <- CT_FOCS[chrom(anchorTwo(CT_FOCS))!="chrY"]













# getting gene coordinates
tss <- readRDS("/data/akalin/Projects/AAkalin_Catalog_RI/final_Version/Data/GENCODE/GENECODE_TSSes.rds")


gene.name.matching <- cbind(match(CT_FOCS$gene.name,tss$gene.name),1:length(CT_FOCS))
gene.name.matching <- gene.name.matching[complete.cases(gene.name.matching),]


FOCS.SS <- GInteractions(anchorOne(CT_FOCS[gene.name.matching[,2],]),
            (tss[gene.name.matching[,1],]))

  mcols(FOCS.SS) <- mcols(CT_FOCS[gene.name.matching[,2],])
  
FOCS.SS$name <- tss$gene.id[gene.name.matching[,1]]
FOCS.SS$name2 <- tss$gene.name[gene.name.matching[,1]]


saveRDS(unique(FOCS.SS),"/data/akalin/Projects/AAkalin_Catalog_RI/post_PhD_analyses/Data/CT-FOCS/CT_FOCS_Processed.rds")


```


```{r}

sessionInfo()
```
